---
title: "Lab 3 - Writing Functions 1"
author: "Cameron Kay & Ashley Miller"
date: "5/8/2019"
output: 
  html_document:
    theme: spacelab
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include = FALSE}
# set knitr options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7.5,
                      fig.height = 4.5)

# load packages
library(tidyverse)
library(magrittr)
library(gapminder)
```

### Part A: Foundations

#### 1. Briefly name and describe the three fundamental components of a function. (Ash)

 + Body: The code supplied to the function that produces the output
 
 + Formals: The arguments supplied to the funciton
 
 + Environment: Our global workspace
 

#### 2. Describe three different ways functions can be stored/applied and an example of when you might want to use each version. (Ash)

+ As an object: If you want the function to be saved to your environment for future use, you can save the function as an object. For example, if you create a function to flip the sign of a number (e.g., `function(x) {x * -1}`) you could save it as `flip_sign` and use it throughout your script.

+ Anonymous functions: If the function is for single use within a loop, you could create an anonymous function. The anonymous function will only exist in the loops environment. Using the above example, the function inside `map(my_vector, function(x) x * -1)` will only exist inside `map`.

+ Storing various functions within a list: If you have multiple similar functions or you want to invoke multiple functions in a loop, you could store the functions as a list. For example, if you had two functions--`flip_sign_1 <- function(x) {x * -1}` and `flip_sign_2 <- function(x) {x * -2}`--you could store them in a list (e.g., `flip_sign <- list(flip_sign_1, flip_sign_2)`) and call them as needed (e.g., `flip_sign[[1]]` or `flip_sign[[2]]`). You could also loop through the list of functions and apply each of them to a object (e.g., `map(flip_sign, ~.x(8))`).

## Part B: Applied practice

**The mean is defined as**

$\frac{∑_{i=1}^nx_i}{n}$

**where $x_i$ represents each element in the mean and $n$ represents the total number of elements. In other words, sum and divide by the number included.**

#### 1. Write a function to calculate the mean that removes missing data before conducting the calculation. You may not use base::mean or any similar function that conducts a mean calculation. (Ash)

```{r partb_1_ash}
# create function
my_mean <- function(x) {
  # remove NAs
  z <- na.omit(x)
  # sum the remaining values
  z_sum <- sum(z)
  # find the length of the object
  z_length <- length(z)
  # calculate the mean given the sum and the length
  mean <- z_sum / z_length
  # return the mean
  mean
}

```

#### 2. Test your function to make sure it (a) provides the expected results, and (b) gives identical output to base::mean when na.rm = TRUE. (Ash)

```{r partb_number2_ash}
# test my function on airquality$Ozone
my_mean(airquality$Ozone)

# test the base::mean function on airquality$Ozone
mean(airquality$Ozone, na.rm = TRUE)

# create vector with missing data
test <- c(1, 2, 3, 4, 5, NA, 8)

# test my function on the test vector
my_mean(test)

# calculate the mean of the test vector using base::mean
mean(test, na.rm = TRUE)
```

#### 3. Write a function that takes two vectors of the same length and returns the total number of instances where the value is NA for both vectors. For example, given the following two vectors (Cam)

```{r b3_example}
x <- c(1, NA, NA, 3, 3, 9, NA)
y <- c(NA, 3, NA, 4, NA, NA, NA)
```

**The function should return a value of 2, because the vectors are both NA at the third and seventh locations. Provide at least one additional test that the function works as expected.**

```{r b3_answer}
# create the function
count_double_nas <- function(v1, v2) {
  sum(is.na(v1) * is.na(v2))
}

# test the function
count_double_nas(x, y)

# test the function with new vectors
z <- c(2, 2, 3, NA, NA, NA, 3, 3, 9, NA, NA)
a <- c(7, NA, 8, NA, NA, NA, 8, 8, 1, 2, NA)

count_double_nas(z, a)
```

#### 4. Turn the following into functions. Make sure to give them meaningful names. Test them out to make sure they provide the expected output. (Ash)

```{r b4_example, eval=FALSE}
mean(is.na(airquality$Ozone))

purrr::map_dbl(split(mtcars$cyl, mtcars$cyl), length) / length(mtcars$cyl)

(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / 
  sd(mtcars$mpg, na.rm = TRUE) 
```

**Function 1:**

```{r partb_num4_1_ash}

mean_missing <- function(x) {
   missing <- is.na(x)
   mean_missing <- mean(missing)
   mean_missing
}

mean_missing(airquality$Ozone)

mean(is.na(airquality$Ozone))

```

**Function 2:**

```{r partb_num4_2_ash}

proportion <- function(x, y) {
  z <- split(x[, y], x[, y])
  z_length <- lapply(z, length)
  prop <- map(z_length, ~(.x/length(x[, y])))
  mean_prop <- map_dbl(prop, ~mean(.x))
  mean_prop
}

proportion(mtcars, "cyl")

purrr::map_dbl(split(mtcars$cyl, mtcars$cyl), length) / length(mtcars$cyl)

```

**Function 3:**

```{r partb_num4_3_ash}

standardize <- function(x) {
  z <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE) 
  z
}

standardize(mtcars$mpg)

(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / 
  sd(mtcars$mpg, na.rm = TRUE) 

```

#### 5. Write a function that takes a data frame as its input and returns a data frame with only the numeric columns. (Cam)

```{r b5_answer}
# create the function
num_cols_only <- function(data) {
  # check if data is a dataframe 
  if (!is.data.frame(data)) {
    stop("Data is not of type dataframe.")
  # if the data is a dataframe stop the function
  } else {
    return(data[, map_lgl(gapminder, is.numeric)])
  }
}

# test the function with a dataframe
num_cols_only(gapminder)

# test the function with a non-dataframe
#num_cols_only(8) #commented out because file will not knit if this runs
```

#### 6. Write a function that uses the function you wrote in Question 5, and returns a data frame with the mean and standard deviation of each numeric column. (Cam)

```{r b6_answer}
# create the function
describe_num_cols <- function(data) {
  # identify only the numeric columns
  data_num_cols <- num_cols_only(data)
  # calculte column means and sds
  data_means    <- map_dbl(data_num_cols, mean)
  data_sds      <- map_dbl(data_num_cols, sd)
  # combine the means and sds 
  data.frame(rbind(data_means, data_sds))
}

# test the function with a dataframe
describe_num_cols(gapminder)

# test the function with a non-dataframe
#num_cols_only(8) #commented out because file will not knit if this runs

# ensure the output is a dataframe
is.data.frame(describe_num_cols(gapminder))
```
